### 2021-06-06 paper/parlr_paper_1.tex
first draft of paper

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Parlr: Parallelized Record Linkage in R}

    Brian Kundinger

@@

Record linkage is the task of identifying duplicate records across
multiple data sources. This task is straightforward when provided unique
identifiers or highly reliable information, but becomes difficult when
data sources are plagued with error. Additionally, many record linkage
tasks are large in nature, and many existing methods are too
computational intensive to be used in practice. In this paper, we
propose a Bayesian model for record linkage that is amenable to parallel
computing, hashing techniques and efficient storage mechanisms, that is
scalable to large linkage tasks.

The proposed method is inspired by the ``Beta Record Linkage'' method
proposed by Mauricio Sadinle's 2017 ``Bayesian Estimation of Bipartite
Matching for Record Linkage''. As such, we demonstrate our method
through a comparison of our methods against Sadinle's on the simulation
study he designed and the El Salvador homocide case study he utilized.
We include one additional simulation study to demonstrate the power of
our method on large datasets. In what follows, Section 2 describes
previous work in record linkage that lays the foundation for our method,
Section 3 provides our model and the Gibbs sampler used for posterior
inference, and Section 4 describes strategies used for more efficient
and scalable implementation. Then, Sections 5 and 6 demonstrate the
method through simulations and a case study. Lastly, our discussion
highlights open questions in the record linkage liturature and motivates
future work.

@@

Most record linkage techniques are derived from the seminal 1969 paper
by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
characteristic of their model was to transform the sets of records,
which often contain text data that is difficult to model, into sets of
comparison vectors governed by parameters that can be more easily
estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
records respectively, and if the files share \(F\) fields in common upon
which to base the linkage, the Fellegi and Sunter approach generates a
\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
between each pair of records across datasets. We say \(\gamma_{ij}\) is
the comparison vector for record \(i \in A\) and record \(j \in B\),
with \(\gamma_{ij}^f\) providing their similarity score on the
\(f^{th}\) field. For ease of modeling and computation, we restrict
these similarity scores to be discrete, ordinal variables, and the
construction of these is left to the modeler. It is common to use use
binary 0-1 variables to indicate exact matching, and 0-1-2 variables to
provide an option for partial matching. For text data, we calculate
similarity based on Levenstein distance or some other text similarity
score, and bin these scores to integers for use in the model.

The likelihood in the original Fellegi Sunter model is a mixture model
through which each record pair is independently classified as a match or
nonmatch. This independent classification often leads to sets of matches
that break expectations of transisitivity, and Jaro devised an procedure
to reduce a set of conflicting matches to the mostly likely set of
one-to-one matchings. One key strength of Sadinle's \texttt{BRL} is that
it is garunteed to produce a set of bipartite matchings that respect
one-to-one assumptions--without the need for any post-hoc procedures.
Specifically, in each iteration of his Gibbs sampler, he considers each
record \(j\in B\), removes from consideration the records \(i\in A\)
that have already been matched, and then samples a potential link.
However, accounting for these dependencies throughout the linkage
process is computationally burdensome, leaving \texttt{BRL} only
suitable for small to moderate linkage problems.

@@

Our notation and assumptions closely follow those of Sadinle. Denote two
files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records respectively,
and with records indexed as \(i \in \{1, \ldots, n_A\}\) in \(A\) and
\(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of generality, label
the files such that \(n_A \geq n_B\). We also assume there are no
duplicates within files, only across. For each record pair under
consideration, we generate a comparison vector
\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
where \(F\) is the number of fields used in the linkage and each
\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
indicating the level agreement between the two records on a specified
field.

@@

\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \] The
prior for \(Z_j\) has equal probability of matching to all records
\(i\in A\), and non-matching probability governed by \(\lambda\).
Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds to a prior
belief that nonmatches and matches are equally likely, and a
\(\lambda \sim \text{Beta}(1, \frac{1}{n_A})\) prior corresponds to a
uniform prior on the labeling of \(\mathbf{Z}\).

Here the reader should the relationship between our proposed model and
that of Sadinle 2017. In his model, Sadinle constructs a prior
distribution on the entire \(\mathbf{Z}\) vector, which induces a Gibbs
sampler that strictly enforces one-to-one matching. In particular, this
sampler removes previously matches records from the set of candidate
records when sampling \(Z_j\), creating a dependency that makes the
sampler \emph{inherently serial}. We however use independent priors for
each \(Z_j\), thereby weakening the one-to-one requirement; our sampler
does ensure that each record in \(B\) can be matched to at most one
record in \(A\), but allows for the possiblity that multiple records in
\(B\) match to the same record in \(A\). We explore the ramifications of
this distinction with the El Salvador case study. With this change,
components of the \(\mathbf{Z}\) vector can be computed in parallel,
allowing for significant computational gains. More importantly, since
only the agreement pattern of \(Z_j\) is used for calculations within
the Gibbs sampler, and not the particular record label, we can conduct
this sampling only at the level of the unique agreement patterns,
offering even more computational savings. This will be explained further
in \textcolor{red}{Section XX}.

@@

(Adapt from Sadinle)

Since our Gibbs procedure does not strictly enforce one-to-one matching,
it is possible for the final Bayes estimate to link multiple records in
\(B\) to one record in \(A\). The modeler can either report both such
matches (with their respective posterior match probabilities), or
resolve these conflicts by accepting only the match with highest
posterior probability. Rigorous proof of this resolution procedure is
deferred to future work.

@@

Broadly speaking, we increase our computational efficiency by
recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. Let
\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
\(P\) denote the total number of unique agreement patterns. Note that
\(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound
does not scale with \(n_A\) or \(n_B\). Prior to processing the data, we
identify all \(P\) patterns in \(\mathcal{H}\) and enumerate them
\(h_1, \ldots, h_P\). When the \((i,j)\) pair exhibits the \(p^{th}\)
agreement pattern, we say \((i,j) \in h_p\). Wherever possible, we
conduct calculations over these \(P\) agreement patterns rather than the
\(n_A \times n_B\) record pairs.

@@

The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
comparison vector as vector of length \(F\), with each component
\(\gamma_{ij}^f\) taking on values in \(\{0, \ldots, L_f - 1\}\). To
ease computations, we instead represent the comparison as a
concatenation of \(F\) many binary indicator vectors of lengths \(L_f\).
For example, if \(L_1 = L_2 = 2\) and \(L_3 = 3\), then
\(\gamma_{ij} = (1, 0, 2)\) under the classical framework becomes
\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
bijective transformation that does not change the meaning of the data,
but this representation will ease calculations and posterior updates.

In the classic Fellegi Sunter framework, \(\Gamma\) is a
\(n_A n_B \times F\) matrix, with each row providing the comparison
vector for a different \((i,j)\) pair. We however do not store these
comparison vectors themselves, but instead only a hashed value
corresponding to the agreement pattern of the \((i, j)\) pair. We store
this information in a nested list \(\tilde{\Gamma}\) where the
\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
in \(A\) that share agreement pattern \(p\) with record \(j \in B\). For
each \(p\), we also calculate
\(H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\),
the total instances of agreement pattern \(p\) throughout the data, and
also for each \(j\), we calculate
\(H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}\) the
instances of agreement pattern \(p\) among the comparison vectors
between record \(j \in B\) and each of the \(n_A\) records in \(A\).

For large data, we can partition the two datasets \(A\) and \(B\) into
smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more manageable
computations. On a single machine, we can read-in data sequentially,
conduct hashing, collect results, and delete the original data from
memory before continuing with the next chunk of data. With multiple
cores or multiple machines, this can be done in parallel. Storing this
hashed information still becomes burdensome for large data, but this
hashing method greatly expands the capabilities of the Fellegi-Sunter
framework.

@@

To address this, we propose a new strategy called \emph{storage
efficient indexing} (SEI). In standard indexing, the modeler decides a
certain criteria that they expect all true matching pairs to satisfy,
and a priori label any record pairs that do not meet that criteria as
non-matching. For example, one might only consider pairs with a certain
similarity score on a field deemed to be important (like first name), or
only pairs with exact matching on a specified number of fields. While
generally chosen to be be quite loose, establishing these criteria
requires knowledge of the problem and invites room for human error. We
propose a method of reducing the comparison space (and reducing storage
requirements) without these drawbacks. Note that all records of the same
agreement pattern have the same probability when sampling \(Z_j\).
Therefore we know that records belonging to an \(h_p\) such that
\(H_{p_j}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate, even without
considering the form of the agreement pattern itself. In random
indexing, rather than store all of these unlikely record labels, we
choose to store only a small number \(R\) of them. Posterior
calculations still attribute the appropriate weight to all records
through the summary statistics \(H_{p_j}, H_p^m\), and \(H_p^u\). Rather
than storing \(n_A \times n_B\) record labels, SEI allows us to store at
most \(n_B \times P \times R\) labels, regardless of how large \(n_A\)
is. Coupled with the partitioning strategy for creating comparison
vectors, this greatly expands the size of the linkage tasks we can
undertake.

@@

\underline{Updating $\mathbf{Z}$:} Although sampling \(Z_j\) from a the
full conditional provided earlier is conceptually straightforward, it
becomes computational burdensome when \(n_A\) is larger. The reader can
confirm that sampling a value from a large set of unequal probabilities
becomes difficult in most programming languages. To speed up
computation, we break this sampling step into two simpler steps. First,
we calculate the Fellegi Sunter weight \(w_{h_p}\) associated with each
unique pattern and sample the agreement pattern between \(j\) and its
potential match. Second, we sample the record label uniformly among
records associated with that agreement pattern for that particular
\(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement
pattern between \(j\) and its potential match, and say
\(h(Z_j) = h_{P+1}\) when \(Z_j = n_A + 1\). Then,

\[h\left(Z_j^{(s+1)}\right) | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}} \propto
\begin{cases} 
    w_{h_p}\times H_{p_j}  & p \leq P; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
\end{cases}\]

Lastly, we recognize that all posterior updates are governed by the
agreement patterns of the record pairs rather than the record labels
themselves. Thuse We complete the entire Gibbs procedure at the level of
the \(P\) agreement patterns. After, we can back-fill the records
corresponding to the agreement patterns by sampling uniformly at random
among candidate records stored in \(\tilde{\Gamma}\). Sampling uniformly
is computational simple even for large sets of candidate records, but
this step can also be parallelized when working with large data.

@@

Comparing our method against \texttt{BRL} on these same simulated
datasets, we find that our method only has weakened performance in the
most extreme scenario of very high errors and very high overlap across
files. We use the following measures of accuracy in our analysis:
although using only recall and precision is common in the literature, we
find it useful to have a single metric the gauge overall performance,
and thus also include F measure.

@@

\begin{center}\includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot} \end{center}

@@

To demonstrate speed, we generate comparison vectors from pre-specified
distributions so that we can easily increase the size of the linkage
problem. Distributions are meant to emulate the behavior of similarity
scores across first name, last name, and day, month, and and year of
birth, with 1 indicating an exact match on a particular field and 0
indicating a nonmatch. We simulate these data for different values of
\(n_A\) and \(n_B\), and compare the run-time of \texttt{parlr} against
\texttt{BRL}. Note that the number of unique patterns \(P\) is bounded
above by \(2^5 = 32\), a bound which is consistently attained in the
larger simulations.

We see that at low data size, \texttt{BRL} outperforms, but that
\texttt{parlr} is significantly faster at handling larger data. In
particular, run-time for \texttt{BRL} seems to grow quadratically (or
linearly with the size of both \(A\) and \(B\)) while run-time for
\texttt{parlr} seems to grow linearly (in the size of \(B\)).

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & m & u \\ 
  \hline
fname\_0 & 0.05 & 0.99 \\ 
  fname\_1 & 0.95 & 0.01 \\ 
  lname\_0 & 0.05 & 0.99 \\ 
  lname\_1 & 0.95 & 0.01 \\ 
  day\_0 & 0.05 & 0.97 \\ 
  day\_1 & 0.95 & 0.03 \\ 
  month\_0 & 0.05 & 0.92 \\ 
  month\_1 & 0.95 & 0.08 \\ 
  year\_0 & 0.05 & 0.93 \\ 
  year\_1 & 0.95 & 0.07 \\ 
   \hline
\end{tabular}
\end{table}

\begin{center}\includegraphics[width=23.6in]{../notes/figures/sadinle_speed_plot} \end{center}

The above discussion suggests that for fixed \(n_B\), computation time
should remain mostly constant with growing \(n_A\). Simulation study
suggests that this is true. In the plot below, fixing \(n_B = 500\), we
see linear growth for the run-time under \texttt{BRL} as \(n_A\)
increases, with much more static run-time under \texttt{parlr}. The
slight increases in run-time that we do see are due primarily to the
hashing step, which again can be run in parallel for large data.

\begin{center}\includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB} \end{center}

We note here that \texttt{BRL} is coded in C, which makes for unfair
comparison against \texttt{parlr}, currently only built in R.
Additionally, although \texttt{parlr} is amenable to parallelization,
this simulation was run on a single core. Running \texttt{parlr} in C++
with paralellization for the hashing step and sampling the matching
status of the record pairs should lead to even more drastic results.

@@

the country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualities, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casuality lists. To obtain a more accurate estimate of the
casualties then, we follow Sadinle's 2017 paper and link records of El
Salvadoran civilian casualties from two sources: El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized reports that
had been published throughout the conflict. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized.

There are several challenges with working with such data. Firstly, both
datasets have been automatically digitized, which inherently leads to
some degree of typographical error. Secondly, the CDHES records are all
second hand accounts reported by individuals, which can result in
additional errors. Lastly, the only fields recorded are given name, last
name, date of death, and place of death; it is relatively common for a
parent and child to share the same given name, resulting in
indistinguishable records for two different individuals. This last point
nearly breaks the above mentioned assumption that there are no
duplicates within files, and reveals a key difference between
\texttt{BRL} and our proposed method.

>>>>

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}

    Brian Kundinger, Rebecca Steorts, and Jerome Reiter

@@

Entity resolution (or record linkage) is the task of identifying
duplicate records across multiple data sources. This task is
straightforward when provided unique identifiers or highly reliable
information, but becomes difficult when data sources are plagued with
error. Additionally, many entity resolution tasks are large in nature,
and many existing methods are too computational intensive to be used in
practice. In this paper, we propose a Bayesian model for record linkage
that uses hashing techniques and efficient storage mechanisms, and is
amenable to parallelization, making it scalable to large linkage tasks.

The proposed method is inspired by the \emph{Beta Record Linkage}
(\texttt{BRL}) method proposed in Mauricio Sadinle's 2017 ``Bayesian
Estimation of Bipartite Matching for Record Linkage''. To recognize this
lineage, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
pronounced ``fable''). In what follows, Section 2 describes previous
work in record linkage that lays the foundation for our method, Section
3 provides our model and the Gibbs sampler used for posterior inference,
and Section 4 describes strategies used for more efficient and scalable
implementation. Then, Section 5 demonstrates the accuracy and speed of
out method through two simulation studies, and Section 6 reveals
interesting properties of the method through a case study of El
Salvadoran casuality records. Lastly, our discussion highlights open
questions in the record linkage liturature and motivates future work.

@@

Most record linkage techniques are derived from the seminal 1969 paper
by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
characteristic of their approach was to transform the sets of records,
which often contain text data that is difficult to model, into sets of
comparison vectors governed by parameters that can be more easily
estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
records respectively, and if the files share \(F\) fields in common upon
which to base the linkage, the Fellegi and Sunter approach generates an
\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
between each pair of records across datasets. We say \(\gamma_{ij}\) is
the comparison vector for record \(i \in A\) and record \(j \in B\),
with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
score on the \(f^{th}\) field. For ease of modeling and computation, we
restrict these similarity scores to be discrete, ordinal variables, and
the construction of these is left to the modeler. We adopt the
convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
is common to use use binary 1-2 variables to indicate exact matching,
and 1-2-3 variables to provide an option for partial matching. For text
data, we calculate similarity based on Levenstein distance or some other
text similarity score, and bin these scores to integers for use in the
model. Jaro (1989) provided guidelines for appropriate binning for the
comparison vectors, but details of implementation are generally context
specific.

The likelihood in the original Fellegi Sunter model is a mixture model
through which each record pair is independently classified as a match or
nonmatch. This independent classification often leads to sets of matches
that break expectations of transitivity, which is undesirable. Although
Jaro devised an procedure to reduce a set of conflicting matches to the
mostly likely set of one-to-one matchings, the abundance of false
matches can bias the estimation of other parameters of the model.

One key strength of Sadinle's \texttt{BRL} is that it is garunteed to
produce a set of bipartite matchings that respect one-to-one
assumptions--without the need for any post-hoc procedures. Specifically,
in each iteration of his Gibbs sampler, he considers each record
\(j\in B\), removes from consideration the records \(i\in A\) that have
already been matched, and then samples a potential link. However,
accounting for these dependencies throughout the linkage process is
computationally burdensome, leaving \texttt{BRL} only suitable for small
to moderate linkage problems.

@@

Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
generality, label the files such that \(n_A \geq n_B\). We assume there
are no duplicates within files, only across. For each record pair under
consideration, we generate a comparison vector
\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
where \(F\) is the number of fields used in the linkage and each
\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
indicating the level agreement between the two records on a specified
field.

@@

\[\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda}) \] The
prior for \(Z_j\) has equal probability of matching to all records
\(i\in A\), and non-matching probability governed by \(\lambda\).
Therefore a \(\lambda \sim \text{Beta}(1, 1)\) corresponds to a prior
belief that nonmatches and matches are equally likely, and a
\(\lambda \sim \text{Beta}\left(1, \frac{1}{n_A}\right)\) prior
corresponds to a uniform prior on the labeling of \(\mathbf{Z}\).

Here the reader should the relationship between our proposed model and
that of Sadinle 2017. In his model, Sadinle constructs a prior
distribution on the entire \(\mathbf{Z}\) vector, which induces a Gibbs
sampler that strictly enforces one-to-one matching. In particular, this
sampler removes previously matches records from the set of candidate
records when sampling \(Z_j\), creating a dependency that makes the
sampler \emph{inherently serial}. We however use independent priors for
each \(Z_j\), creating a sampler that is \emph{perfectly
parallelizable}, allowing for significant computational gains. More
importantly, since only the agreement pattern of \(Z_j\) is used for
calculations within the Gibbs sampler, and not the particular record
label, we can conduct this sampling only at the level of the unique
agreement patterns, offering even more computational savings. In doing
so however, we thereby weaken the one-to-one requirement from
\texttt{BRL}; our sampler does ensure that each record in \(B\) can be
matched to at most one record in \(A\), but allows for the possiblity
that multiple records in \(B\) match to the same record in \(A\). We
explore the ramifications of this distinction with the El Salvador case
study.

@@

We calculate a Bayes estimate \(\hat{\mathbf{Z}}\) for the linkage
parameter \(\mathbf{Z}\) by assigning different positive losses to
different types of errors, and minimizing posterior expected loss. We
adopt the loss function proposed in Sadinle 2017, in which
\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
option to leave the matching undetermined by the model. Specifically, we
minimize the quantity
\(L(\hat{\mathbf{Z}}, \mathbf{Z}) = \sum_{j=1}^{n_B} L(\hat{Z_j}, Z_j)\)
where

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
    0  & \text{if } Z_j = \hat{Z_j}; \\
    \theta_R,  & \text{if } \hat{Z_j} = R; \\
    \theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
    \theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
    \theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j} ; \\
\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
on the linkage status, \(\theta_{10}\) is the loss from a false
non-match, \(\theta_{01}\) is the loss from a false match, and
\(\theta_{11}\) is the loss from the special case of a false match in
which the record has a true match other than the one estimated by the
model. This loss function leads to closed form decision rules for
minimizing posterior expected loss. In this paper, we adopt losses
\(\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2\),
inducing the intuitive decision rule

\[\hat{Z}_j =\begin{cases} 
    i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
    0,  & \text{otherwise} ; \\
\end{cases}\]

For a more indepth explanation of this function and the induced Bayes
estimate, see Sadinle 2017.

Since our Gibbs procedure does not strictly enforce one-to-one matching,
it is possible for the final Bayes estimate to link multiple records in
\(B\) to one record in \(A\). The modeler can either report both such
matches (with their respective posterior match probabilities), or
resolve these conflicts by accepting only the match with highest
posterior probability. Such a resolution procedure indeed is equivalent
to minimizing posterior risk under the restriction of one-to-one
matching, and as thus theoretically justified within the Bayesian
framework.

@@

Broadly speaking, we increase our computational efficiency by
recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. Let
\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
\(P = |\mathcal{H}|\) denote the total number of unique agreement
patterns. Note that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
that this bound does not scale with \(n_A\) or \(n_B\). Prior to
processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
and enumerate them \(h_1, \ldots, h_P\). When the \((i,j)\) pair
exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
Wherever possible, we conduct calculations over these \(P\) agreement
patterns rather than the \(n_A \times n_B\) record pairs.

@@

The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
comparison vector as vector of length \(F\), with each component
\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). To ease
computations, we instead represent the comparison as a concatenation of
\(F\) many binary indicator vectors of lengths \(L_f\). For example, if
\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
under the classical framework becomes
\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
bijective transformation that does not change the meaning of the data,
but this representation will ease calculations and posterior updates.

In the classic Fellegi Sunter framework, \(\Gamma\) is a
\(n_A n_B \times F\) matrix, with each row providing the comparison
vector for a different \((i,j)\) pair. In contrast, we do not store
these comparison vectors themselves, but instead only a hashed value
corresponding to the agreement pattern of the \((i, j)\) pair. We store
this information in a nested list \(\tilde{\Gamma}\) where the
\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
in \(A\) that share agreement pattern \(p\) with record \(j \in B\). For
each \(p\), we also calculate
\(H_p = \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} \mathbf{1}_{(i,j) \in h_p}\),
the total instances of agreement pattern \(p\) throughout the data, and
also for each \(j\), we calculate
\(H_{p_j} = \sum_{i=1}^{n_A} \mathbf{1}_{{(i,j) \in h_p}}\) the
instances of agreement pattern \(p\) among the comparison vectors
between record \(j \in B\) and each of the \(n_A\) records in \(A\).

@@

To address this, we propose a new strategy called \emph{storage
efficient indexing} (SEI). In standard indexing, the modeler decides a
certain criteria that they expect all true matching pairs to satisfy,
and a priori label any record pairs that do not meet that criteria as
non-matching. For example, one might only consider pairs with a certain
similarity score on a field deemed to be important (like first name), or
only pairs with exact matching on a specified number of fields. While
generally chosen to be be quite loose, establishing these criteria
requires knowledge of the problem and invites room for human error. We
propose a method of reducing the comparison space (and reducing storage
requirements) without these drawbacks. Note that all records of the same
agreement pattern have the same probability when sampling \(Z_j\).
Therefore we know that records belonging to an \(h_p\) such that
\(H_{p_j}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate, even without
considering the form of the agreement pattern itself.

In SEI, rather than store all of these unlikely record labels, we choose
to store only a small number \(R\) of them. Posterior calculations still
attribute the appropriate weight to all records through the summary
statistics \(H_p\), and \(H_{p_j}\). Rather than storing
\(n_A \times n_B\) record labels, SEI allows us to store at most
\(n_B \times P \times R\) labels, regardless of how large \(n_A\) is.

Laslty, for large data, we can partition the two datasets \(A\) and
\(B\) into smaller blocks \(\{A_m\}\) and \(\{B_m\}\) for more
manageable computations. On a single machine, we can read-in data
sequentially, conduct hashing, compress information through SEI, and
delete the original data from memory before continuing with the next
chunk of data. With multiple cores or multiple machines, this can be
done in parallel. Thus the combination of hashing, SEI, and partitioning
allows us to conduct linkage tasks over much larger datasets.

@@

\underline{Updating $\mathbf{Z}$:} Although sampling \(Z_j\) from a the
full conditional provided earlier is conceptually straightforward, it
becomes computational burdensome when \(n_A\) is larger. The reader can
confirm that sampling a value from a large set of unequal probabilities
becomes difficult in most programming languages. To speed up
computation, we break this sampling step into two simpler steps. First,
we calculate the Fellegi Sunter weight \(w_{p}\) associated with each
unique pattern and sample the agreement pattern between \(j\) and its
potential match. Second, we sample the record label uniformly among
records associated with that agreement pattern for that particular
\(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement
pattern between \(j\) and its potential match, and say
\(h(Z_j) = h_{P+1}\) when \(Z_j = n_A + 1\). Then,

\[P\left(h\left(Z_j^{(s+1)}\right) = p | \mathbf{m}, \mathbf{u}, \mathbf{Z^{(s)}}\right) \propto
\begin{cases} 
    w_{p}\times H_{p_j}  & p \leq P; \\
     n_A \frac{n_B - D + \beta_{\lambda}}{D + \alpha_{\lambda}} &   p = P + 1 \\
\end{cases}\]
\[P\left(Z_j^{(s+1)} = i\bigg|\;\; h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
\frac{1}{H_{p_j}} & (i, j) \in h_p \\
0 & \text{otherwise} \\
\end{cases}\]

Lastly, we recognize that all posterior updates are governed by the
agreement patterns of the record pairs rather than the record labels
themselves. Thus we complete the entire Gibbs procedure first at the
level of the \(P\) agreement patterns with the first equation above.
After, we can back-fill the records corresponding to the agreement
patterns through the second equation. Sampling uniformly is
computational simple even for large sets of candidate records, but this
step can also be parallelized when working with large data.

@@

Comparing our method against \texttt{BRL} on these same simulated
datasets, we find that our method only has weakened performance in the
most extreme scenario of very high errors and very high overlap across
files. We use the following measures of accuracy in our analysis:
although using only recall and precision is common in the literature, we
find it useful to have a single metric to gauge overall performance, and
thus also include F measure.

@@

\begin{figure}

{\centering \includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot}

}

\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017. For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:unnamed-chunk-1}
\end{figure}

@@

To demonstrate speed, we generate comparison vectors from pre-specified
distributions so that we can easily increase the size of the linkage
problem. Distributions are meant to emulate the behavior of similarity
scores across first name, last name, and day, month. For simplicity, we
consider only exact matching, so a vector (1, 0) corresponds to
agreeement and (0, 1) to disagreement. We simulate these data for
different values of \(n_A\) and \(n_B\), and compare the run-time of
\texttt{fabl} against \texttt{BRL}. Note that the number of unique
patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is
consistently attained in the larger simulations.

We see that at low data size, \texttt{BRL} outperforms, but that
\texttt{fabl} is significantly faster at handling larger data. In
particular, run-time for \texttt{BRL} seems to grow quadratically (or
linearly with the size of both \(A\) and \(B\)) while run-time for
\texttt{fabl} seems to grow linearly (in the size of \(B\)).

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & m & u \\ 
  \hline
  fname & (0.95, 0.05) & (0.01, 0.99) \\ 
  lname & (0.95, 0.05) & (0.01, 0.99) \\ 
  day & (0.95, 0.05) & (0.03, 0.97) \\ 
  month & (0.95, 0.05) & (0.08, 0.92) \\ 
  year & (0.95, 0.05) & (0.07, 0.93) \\ 
   \hline
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=17.75in]{../notes/figures/sadinle_speed_plot2}

}

\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:unnamed-chunk-2}
\end{figure}

The above discussion suggests that for fixed \(n_B\), computation time
should remain mostly constant with growing \(n_A\). Simulation study
suggests that this is true. In the plot below, fixing \(n_B = 500\), we
see linear growth for the run-time under \texttt{BRL} as \(n_A\)
increases, with much more static run-time under \texttt{fabl}. The
slight increases in run-time that we do see are due primarily to the
hashing step, which again can be run in parallel for large data.

\begin{figure}

{\centering \includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB}

}

\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:unnamed-chunk-3}
\end{figure}

We note here that \texttt{BRL} is coded in C, which makes for unfair
comparison against \texttt{fabl}, currently only built in R.
Additionally, although \texttt{fabl} is amenable to parallelization,
this simulation was run on a single core. Running \texttt{fabl} in C++
with paralellization for the hashing step and sampling the matching
status of the record pairs should lead to even more computational gains.

@@

The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualities, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casuality lists. To obtain a more accurate estimate of the
casualties then, we follow Sadinle's 2017 paper and link records of El
Salvadoran civilian casualties from two sources: El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized reports that
had been published throughout the conflict. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized.

There are several challenges with working with such data. Firstly, both
datasets have been automatically digitized, which inherently leads to
some degree of typographical error. Secondly, the CDHES records are all
second hand accounts reported by individuals, which can result in
additional errors. Lastly, the only fields recorded are given name, last
name, date of death, and place of death; it is relatively common for a
parent and child to share the same given name, resulting in
indistinguishable records for two different individuals. This last point
nearly breaks the earlier mentioned assumption that there are no
duplicates within files, and reveals a key difference between
\texttt{BRL} and our proposed method.

We only utilized records with nonmissing entries for given and last
name, results in \(n_A = 4420\) files in CHDES and \(n_B = 1323\) files
in ERTL. The names were standardized to account for common misspellings
in the Spanish language, and then compared using a modified Levenstein
distance to account for the fact that second names are often ommited.
Place of birth is recorded by municipality and department within that
municipality; however, since department was missing in 95\% of records
in CHDES and 80\% of records in ERTL, we excluded it from our linkage
process. Thus we conduct linkage using given name, last name,
municipality, and day, month, and year of death. We again use flat
priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters.

\begin{figure}

{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/m_posterior_smallP}

}
